---
title: "GSE111014"
output: html_notebook
---

```{r}
library(dplyr)
library(patchwork)
library(Seurat)
```


```{r}
cll <- CreateSeuratObject(Read10X("../data/GSE111014/"), project = "cll", min.cells = 3, min.features = 200)
cll
```

```{r}
cll[['percent.mt']] <- PercentageFeatureSet(cll, pattern = "^MT-")
VlnPlot(cll, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
plot1 <- FeatureScatter(cll, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cll, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

Specifically, single-cell profiles with less than 200 detected genes (indicative of no cell in the droplet), more than 3000 detected genes (indicative of cell duplicates), or more than 15% of UMIs stemming from mitochondrial genes were discarded.

```{r}
cll <- subset(cll, subset = nFeature_RNA > 200 & nFeature_RNA < 3000 & percent.mt < 15)
cll
```
Read counts were normalized dividing by the total UMI count in each cell, multiplied by a factor of 10,000, and log transformed.
```{r}
cll <- NormalizeData(cll, normalization.method = "LogNormalize", scale.factor = 10000)
```

The number of UMIs per cell and the percent of mitochondrial reads per cell were then regressed out using Seurat’s standard analysis pipeline.
```{r}
cll <- ScaleData(cll, vars.to.regress = c("percent.mt", "nCount_RNA"), features = rownames(cll))
```

