---
title: "GSE118614 10x"
output: html_notebook
---

This is an analysis for the dataset [GSE118614](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE118614) made available by Clark BS et al. in their paper [_Single-cell RNA-Seq Analysis of Retinal Development Identifies NFI Factors as Regulating Mitotic Exit and Late-Born Cell Specification_](https://www.sciencedirect.com/science/article/abs/pii/S0896627319303459).



```{r}
rm(list=ls())
gc()

library(Matrix)
library(ggplot2)
library(Rtsne)
library(matrixStats)
```

# Overview of the dataset
We first overview the dataset.

```{r}
barcode.path <- "../data/raw/GSE118614/GSE118614_10x_barcodes.tsv.gz"
features.path <- "../data/raw/GSE118614/GSE118614_10x_genes.tsv.gz"
matrix.path <- "../data/raw/GSE118614/GSE118614_10x_aggregate.mtx.gz"
mat <- readMM(file = matrix.path)
feature.names = read.delim(features.path, 
                           header = TRUE,
                           stringsAsFactors = FALSE)
barcode.names = read.delim(barcode.path, 
                           header = TRUE,
                           stringsAsFactors = FALSE)
```

## Cell features

```{r}
barcode.names
```


```{r}
table(barcode.names$age)
table(barcode.names$sample)
table(barcode.names$umap2_CellType)
table(barcode.names$age, barcode.names$umap2_CellType)
table(barcode.names$sample, barcode.names$umap2_CellType)
table(barcode.names$used_for_pseudotime)
```

## Downsampling 
Processing so many cells at the same time is computationally prohibitive. We downsample each age to 500 cells. For those with replicates, we pick from each sample evenly.

```{r}
mask <- barcode.names$used_for_pseudotime == TRUE
mat <- mat[mask, ]
barcode.names <- barcode.names[mask, ]
gc()

#saveRDS(list(mat, barcode.names, feature.names), '../data/gse118614_10x_30k.RDS')
```

```{r}
table(barcode.names$sample)
```


```{r}
set.seed(0)

mask = rep(FALSE, dim(mat)[1])
plate = c('E11', 'E12_rep1', 'E14_rep1', 'E14_rep2', 'E16', 'E18_rep2', 'E18_rep3', 'P0', 'P14', 'P2_rep2', 'P2_rep3', 'P5', 'P8_rep1', 'P8_rep2')
num = c(400, 400, 200, 200, 400, 200, 200, 400, 400, 200, 200, 400, 200, 200)

mask = c()

for (i in 1:length(plate)){
  mask <- c(mask, 
            sample((1:dim(mat)[1])[barcode.names$sample == plate[i]], num[i]))
}

length(mask)
```

```{r}
mat <- mat[mask, ]
barcode.names <- barcode.names[mask, ]
gc()
```

```{r}
mat <- as.matrix(mat)
```

```{r}
table(barcode.names$age)
table(barcode.names$sample)
table(barcode.names$umap2_CellType)
table(barcode.names$age, barcode.names$umap2_CellType)
table(barcode.names$sample, barcode.names$umap2_CellType)
```

```{r}
head(feature.names)
```

```{r}
head(barcode.names)
```

```{r}
colnames(mat) <- feature.names$id
rownames(mat) <- rownames(barcode.names)
```


```{r}
source("../R/varactor.R")
```

```{r}
obj <- Varactor$new(data = list(expr=t(mat)), 
                    labels = list(expr = list(plate = barcode.names$sample, type=barcode.names$umap2_CellType, age=barcode.names$age)), 
                    what = 'raw')

obj$normalize()
#obj$normalized <- obj$raw
obj$combine(sd_threshold = .1, cv_threshold = 0.1)
dim(obj$combined)
obj$reduce(reduce_dim = 50)
obj$clean()
```

```{r}
set.seed(0)
obj$define_metric("primary", "euclidean")$measure("primary")$embed("primary", "tsne")
obj$plot("primary", "embedding", "type")
obj$plot("primary", "embedding", "plate")
obj$plot("primary", "embedding", "age")
```


```{r}
set.seed(0)
obj$define_metric("alternative", "davidson", strata = 'plate')$measure("alternative")$embed("alternative", "tsne")
obj$plot("alternative", "embedding", "type")
obj$plot("alternative", "embedding", "plate")
obj$plot("alternative", "embedding", "age")

obj$labels$type2 <- obj$labels$type
obj$labels$type2[obj$labels$type2 %in% c('Early RPCs', 'Intermediate RPCs', 'Late RPCs')] = 'RPCs'
obj$plot("alternative", "embedding", "type2")
```

```{r}
set.seed(0)
obj$embed("primary", "umap")
obj$plot("primary", "embedding", "type")
obj$plot("primary", "embedding", "plate")
obj$plot("primary", "embedding", "age")
```

```{r}
set.seed(0)
obj$embed("alternative", "umap")
obj$plot("alternative", "embedding", "type")
obj$plot("alternative", "embedding", "plate")
obj$plot("alternative", "embedding", "age")

obj$labels$type2 <- obj$labels$type
obj$labels$type2[obj$labels$type2 %in% c('Early RPCs', 'Intermediate RPCs', 'Late RPCs')] = 'RPCs'
obj$plot("alternative", "embedding", "type2")
```

